Parametric controls on vegetation responses to biogeochemical forcing in the CLM5

### 2019-08-30 main_revision.tex

\item The control model state, and it's responses to CO$_{2}$ and nitrogen addition. are sensitive to parameter choice, and parameter responses further vary with pre-fertilization water and nitrogen limitation. \\

\item Parameters controlling nitrogen fixation rates dominate fertilization responses. Many other fertilization responses are modified indirectly, via changes in the pre-fertilization state. 
\end{keypoints}

@@

Low values of \emph{stem\_leaf} ratio resulted in higher LAI at all sites, as well as declines in total vegetation carbon (.   The impact of increasing leaf allocation is to invoke large increases in LAI and NPP. Lower stem allocation and higher LAI also increase N demand, reducing $N_{leaf}$ and increase PNU at Niwot Ridge and Metolius in Figure \ref{btran state}).

Increasing $N_{costs}$ increases the respiration losses associated with nitrogen uptake.  All sites showed a moderate decline in LAI, NPP, GPP and vegetation carbon when costs were high, but lesser impacts when costs were reduced (Figures \ref{NR1 state}, \ref{CAX state} \ref{HVF state}, \ref{MET state}). In contrast to the allocation parameters, $N_{costs}$ does not appear to be a first-order control over the pre-fertilization system state.

Increasing \emph{fracfixers} affects the maximum amount of carbon that plants can pay towards symbiotic nitrogen uptake (N fixation). For all sites and variables here, it has a relatively moderate and positive impact on the control state. As is the case for $N_{costs}$, \emph{fracfixers} does not appear to be a first-order control over the pre-fertilization system state.

Changing \emph{leafcn} over one standard deviation from the mean TRY values, had substantial impacts on all the output variables at all sites. Higher \emph{leafcn} values decrease $N_{leaf}$, decreasing photosynthetic capacity and GPP.  In particular, at MET and HVF the higher \emph{leafcn} values had $N_{leaf}$ and photosynthetic capacities so low that GPP was almost zero (Figures \ref{HVF state}, \ref{MET state}).  GPP continues to increase almost linearly with declining \emph{leafcn}  (increasing $N_{leaf}$) at all sites except HVF, indicating a lack of saturation of photosynthesis with respect to photosynthetic capacity.  This is consistent with the analysis by \cite{lawrence2018} who find that V${_c,max}$ numbers in CLM5 might be low compared to observations.  For NPP, respiration costs of modified $N_{leaf}$ and altered PNU (Figure \ref{btran state}) buffered the larger changes in GPP. At CAX, for example, the lowest \emph{leafcn} had a GPP 225 gC/m$^{2}$/y higher, while the corresponding NPP increase was only 52 gC/$m^{2}$/y.

@@

For CO$_{2}$ fertilization at CAX the initial stimulation of GPP ($\sim$1.37) lead to an almost 50\% increase in LAI. At this site, water limitations (lower $B_{tran,mn}$), prevented any further increase in GPP.

 At MET, in contrast, initial GPP stimulation was lower (26\%) on account of the lesser water stress. This resulted in slow but prolonged LAI increase, given the lower overall carbon turnover at this colder site. The 50\% increase in carbon at 80 years had, however, a much larger impact on GPP and NPP giving a stimulation of 53-69\% and 49-71\% respectively, (varying through the cycling climate forcing data). Thus, the overall impact of fertilization appears much greater than is typically reported in elevated CO$_{2}$ experiments, but \cite{king2005} report both large response to $sim$530 ppm CO$_{2}$ in Aspen stands, with increases in LAI and biomass of around 60\%.

 At HVF, the initial stimulation in the first year was lower (25\%) potentially on account of the higher carboxylation capacity at this site (HVF has a smaller response to decreasing higher leaf N ratios, Fig. \ref{HVF state}). NWR had moderate (25\%) increases in initial GPP and LAI ($sim$30\%).  At all sites, the fraction of NPP spent on N uptake increased, but most dramatically at Metolius, where initial N expenditure was lowest initially on account of low growth rates.

N addition most strongly affects carbon expenditure on N uptake, and thus NPP, rather than GPP directly. At NWR an initial NPP stimulation of 1.3 for NPP was magnified by subsequent increases in LAI to nearly 1.8 times the control value, causing a stimulation of 1.7 in NPP, reflected the high N limitation at this site in the pre-fertilization state. None of the other sites exhibited such a dramatic change. In particular, for CAX, which was heavily water limited in the first instance, adding nitrogen only resulting in stimulation in LAI, GPP and NPP of around 10\%.

The time-dependence of the responses generates an interesting issue for interpretation. The first year responses are illustrative of the short-term immediate impacts of change, whereas the longer-term responses are heavily dominated by the feedback responses. Further, depiction of absolute vs. relative differences can serve to highlight different aspects of the model response. Here we report values from 15-20 years after fertilization began, to be consistent with the methodology used by \cite{wieder2019} in their analysis of global fertilization of the default model, and to capture the dynamics of the whole system feedbacks captured in Fig. \ref{timescales}.  Thus, figures \ref{GPP_CN} and \ref{LAI_CN} illustrate the responses to 15-20 years of CO$_{2}$ and N  fertilization simultaneously, for GPP and LAI.  Note that each figure is a different scale to best illustrate the parametric impacts.  Corresponding figures for NPP, vegetation carbon and $N_{area}$ are in the supplemental information (SI Figures 1-3). Figures for each site showing the individual N and CO$_{2}$ responses are also included in the SI (Figures 3 to 10).

@@

The impact of \emph{slatop} on fertilization was different at every site (figs. \ref{GPP_CN}, figures \ref{LAI_CN}). For MET, \emph{slatop} was the most important control over N responses, with higher values allowing a greater LAI/GPP feedback in response to the N fertilization. At NWR, the opposite behaviour was found, with high SLA reducing the impact of N in cases where the pre-fertilization LAI meant that a saturating response of GPP was found. At HVF, higher values of \emph{slatop} slightly reduced both CO$_{2}$ and N fertilization for the same reason (although the range of variation at HVF is only 5\% between the highest and lowest values, as opposed to 40\% at NWR).

\emph{froot\_leaf} had contrasting impacts at different sites.  At MET,  low values increased the CO$_{2}$ fertilization, probably due to the increased drought stress under those conditions. At HVF, low \emph{froot\_leaf} decreased both CO$_{2}$ and N fertilization on account of higher initial LAI, mirroring the respnse to \emph{slatop}.

Low \emph{stem\_leaf} ratio caused lower responses to N fertilization CAX, NWR and HVF for GPP and LAI at  (Figures \ref{GPP_CN} and \ref{LAI_CN}), again mirroring the NPP-LAI non-linearity issue.  At NWR,  N limitation restricted CO$_{2}$ responses, but at CAX and NWR the same type of response was seen.  At MET, lower \emph{stem\_leaf} allowed greater responses, plausibly from increasing the speed of the NPP/LAI feedback.

Higher $N_{costs}$ increase the responsiveness of the system to N at CAX, NWT and HVF.  The magnitude of the impact on N fertilization ism as might be expected, related to the extent of N limitation in the pre-fertilization state (Figure \ref{btran state}).   Very low values of $N_{costs}$ appeared to reduce the responsiveness of the system to CO$_{2}$ fertilization, particularly at CAX and HVF **.

\emph{frac\_fixers} (the maximum fraction of NPP available for N fixation) had a particularly dominant role over the model response to step changes in N and to CO$_{2}$. For GPP and LAI, all sites displayed a 'trade-off', whereby high \emph{frac\_fixers} systems responded strongly to CO$_{2}$ but not N, and low \emph{frac\_fixers} systems responded strongly to N but not CO$_{2}$ (Figure \ref{GPP_CN}, \ref{LAI_CN}).   At MET,  the changing fixation capacity had no impact on the response of GPP or LAI to N fertilization.  Only \emph{slatop} and \emph{stem\_leaf} had large impacts on N fertilization for this site (which had the smallest PNU in the first instance), suggesting fundamental metabolic limitations to carbon economics and growth (e.g. a short growing season).

The impact of \emph{leafcn} on fertilization responses was not as dominant a control over fertilization responses as it was over the system state. Complex responses were illustrated, reflecting the differential costs and benefits of Ni on the system.  For N fertilization, at CAX and NWR, lower values of \emph{leafcn} had smaller responses to N fertilization, reflecting the higher initial LAI of these configurations and the non-linear GPP feedback discusses earlier.   At CAX, the highest \emph{leafcn} had smaller CO$_{2}$ responses. (Figure \ref{btran state} illustrates that this configuration had less water stress, given that both the lower LAI and lower photosynthetic capacity should lead to smaller stomatal conductances and evapotranspiration. and thus had less water stress to be alleviated by the high CO$_{2}$.

Surprisingly, given the importance of \emph{medlyn\_slope} on stomatal responses to CO$_{2}$, we did not find a widespread impact of the Medlyn parameter on the fertilization response. Earlier studies using a less constrained range of \emph{medlyn\_slope} values indicated that it can have a potentially large impact on system state, (results not shown) but that the fertilization response is typically limited. The linear response of the stomatal conductance term to $CO_{2}$ mean that the direct impact of \emph{medlyn\_slope} is likely to be restricted to changes in the pre-fertilization state.

The impact of \emph{cn\_flex\_c} was small compared to the more dominant parameters. Higher values of \emph{cn\_flex\_c} spend less carbon on N uptake when tissue N is lower than the target amount, allowing N ratios to vary more. Impacts of this parameter are thus manifested through their impact on $N_{leaf}$ (Figure SI **).  Smaller values of \emph{cn\_flex\_c} corresponded in all cases to smaller variations in $N_{leaf}$ (values closer to 1.0) under both CO$_{2}$ and N fertilization (For N deposition, $N_{leaf}$ always increases and for CO$_{2}$ it always declines.) Lower values of \emph{cn\_flex\_c} also corresponded to larger CO$_{2}$ stimulation of GPP, illustrating the direct impact of this parameter, as leaves maintain higher photosynthetic capacity under more N limiting conditions. The impact on NPP and LAI was less clear, given the corresponding impacts of higher respiration both for maintenance and for N uptake. At CAX, NWR and HVF,  the highest values of \emph{cn\_flex\_c} \emph{also} had a higher response to CO$_{2}$ than the default. The reasons for this discontinuity are unclear, however, the impact of \emph{cn\_flex\_c} in the CLM5 code is subject to at least two 'caps', (whereby plants cannot spend less than 0.5 or more than 1.0 of the carbon that would retain the target LAI), potentially leading to complex feedbacks once these caps are reached.

\subsection{Low leaf area investigations}
In general, we found that LAI predicted by the default CLM5 were often low  compared to published estimates for these sites. (CAX; 3.0 (modeled) vs. 5.3 (observed) \cite{fisher2007},  NWR; 1.7 vs. 4.2 \cite{bowling2009}, HVF; 2.6 vs. 3.5 \cite{williams1996}  MET; 0.5 vs. 1.5 (\cite{spadavecchia2011}). This is a somewhat surprising result, given the generally good agreement of globally-driven model output with LAI reported by \cite{lawrence2018}, and may plausibly reflect differences between the use of reanalysis products used to drive the model and site-specific meteorological data.

To interrogate this further, we ran a separate set of simulations with higher allocation to leaves (generated from lower stem allocation parameters). These simulations indicated that, with the resulting higher (and more consistent with observed) LAI values, the model was less sensitive to both CO$_{2}$ and N deposition. For example, at HVF, the high leaf allocation simulation (LAI of 4.6, vs 2.6 for the default parameters) had a lower CO$_{2}$ response of  (1.18, vs 1.43 for the control)  and a lower N response as well (1.20 vs 1.26 for the control). Similar results were found at all other sites.

\section{Discussion}
The overall responses of the land surface components of Earth system models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP6 model inter-comparison projects and associated activities (\cite{meehl2014}, \cite{eyring2016}). Throughout the CMIP activities, the carbon-cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space. Interpretation of these simulations should be informed by solid comprehension of which model properties dominate the size of the feedback strength both within and between models, and we intend these illustrative simulations to provide initial guidance of this process.

At the time of writing, we do not yet know the carbon-climate feedback strength of the default version of CLM5/CESM2. In offline simulations forced with a data atmosphere, CLM5 shows a stronger response to elevated CO$_{2}$ and a weaker response to N fertilization than previous versions of the model (CLM4 and CLM4.5, Wieder et al. 2018). Further, the impact of both types of fertilization was found, in all versions of the model, to be highly variable with PFT (from 0\% to 110\% for N, -10\% to +55\% for CO$_{2}$). In this study we illustrate how variations in the pre-fertilization state, water and N limitation, and parametric variation combine to generate variation in the responses to forcing. While we set out primarily to to determine responses to model parameters, findings concerning the impact of model state on forcing sensitivity will have important implications for the interpretation of simulations with the default model.

@@

Aside from the straightforward direct impacts of the fixation rate on model predictions, however, the model output space was notably complex. For example,  in some cases the sign of the impact of individual parameters changes between sites, while other parameters can demonstrate non-linear and counter-intuitive impacts as the result of numerous interacting feedbacks.  To summarize our findings, we note a variety of key outcomes of our analysis on the factors controlling responses to CO$_{2}$ and N fertilization in the CLM5, including:

@@

2. The degree of water or nutrient limitation in the initial state is also a dominant factor over the model responses to CO$_{2}$ fertilization and N deposition. Once again, the direct impact of parameter perturbation is frequently masked by how changes affect the initial limiting conditions. Examples of this include how low \emph{froot\_leaf} increases water limitation at NWR, increasing responsiveness to CO$_{2}$ and how high \emph{leafcn} reduces water stress and thus decreases the CO$_{2}$ response at CAX.

3. Other parameters with a direct impact on fertilization responses included those related to the acquisition of nitrogen (\emph{N\_costs}) and the stoichiometric flexibility of the system (\emph{cn\_flex\_c}) illustrate the sensitivity of the model to relatively unconstrained processes within the nitrogen cycle.

4. Many of the parameters that are dominant over the initial state (\emph{slatop}, \emph{stem\_leaf} \emph{leafcn}) either have limited impacts on fertilization responses, or have impacts that are largely traceable to their impacts on the initial model state. In contrast, numerous parameters that more directly and uniformly impact fertilization responses, (\emph{frac\_fixers} \emph{N\_costs}) are less important for the initial model state. We thus suggest that model calibration efforts that only take into account parameters that affect the control model state, are unlikely to constrain the major axes of responses to CO$_{2}$ and N fertilization.

5. The results can in general be interpreted as being in line with the expected dynamics of the model, given sufficient output data and consideration of the dynamics of the individual sites. Given the complexity of the model, this helps generate confidence that the model outcomes are derived from the intended process mechanisms, and are not subject to significant deviations from the intent of the model structure (as opposed to .

@@

\subsection{Model improvements highlighted by this analysis}
While there are many potential improvements that could be made to any land surface model, this particular assessment of the parameter space of the CLM5 in this study revealed several dynamics which highlight parts of the representation of biogeochemistry which might be improved in future versions.  
Firstly, the response of NPP to increasing \emph{slatop} was often negative, despite increasing GPP, as the fixed-fraction allocation scheme in the model will generate leaves that respire more than they assimilate under very high productivity conditions. This highlights the need for either more realistic tissue allocation, or for a scheme which optimizes leaf allocation to prevent leaf layers in negative carbon balance (\cite{fisher2010}).

@@

Thirdly, requiring focus is the dominance of the 'fraction of fixers' over fertilization responses. \emph{frac\_fixers} is a proxy for the community level composition of nitrogen fixing plants. Technically, because plant density is not represented in the CLM5, this parameter represents the fraction of net assimilated carbon ($gpp$ - $r_{auto}$) that can be used for symbiotic N fixation. Ideally, this feature of the model would be prognostic, and the fraction of N fixing plants would respond to changes in environmental conditions and the relative competitive abilities of N fixing plants. In general, Nitrogen fixers are known to primarily exist in early-successional parts of ecosystems (\cite{vitousek1989}), and therefore representation of their ecological dynamics would likely include a representation of vegetation demographics and succession (\cite{fisher2018vegetation}, \cite{trugman2016climate}).

@@

\subsection{Applicability of results to other land surface models}
Our finding that Nitrogen fixation plays a dominant role over responses to $CO_{2}$ fertilization concurs with earlier studies. Notably, \cite{meyerholt2015} investigate six alternative models of N fixation implemented within the O-CN model (\cite{zaehle2010}). They emphasize that the formulation which, as in CLM5, switches N sources depending on their relative costs, represent a promising means by which N fixation might be simulated. Interestingly, their model also includes a maximum rate of fixation per unit (root) biomass, but analysis to it's sensitivity is not shown. Further, given the difficulty of matching observations of complex N cycle processes with more abstract 'big leaf' models such as CLM5, and the ecological complexities governing the abundance their abundance, \cite{meyerholt2015} also advocate for the dynamic simulation of the distribution of N fixing plants within ESMs, as opposed to the assumption that maximum N fixing capacity if static in space and time.

Given that the response of GPP is almost by definition non-linear with respect to LAI, (given the saturating properties of light absorbtion in increasingly dense canopies) we anticipate that the impacts of initial state will be similar in most land surface models. While it is perhaps not a novel point, the degree of water limitation should act to modulate CO$_{2}$ fertilization while N fertilization will of course be stronger in areas that are more N limited in the first instance.

@@

\subsection{Caveats and future directions}
Many kinds of further analyses are of course possible, including ramping transient simulations,  including global feedback analyses with alternative parameterizations, calibration of critical parameters against manipulation experiments, wider assessment against greater ranges of flux tower data, and tighter constrains of the parameter ranges used. We hope that this initial assessment of the CLM5 parameter space will provide a starting point for such activities.

Here we investigate the parameter space of the model using step-change experiments. These numerical experiments have the advantage that they can be informed by field experiments (\cite{wieder2019}, however, such experiments necessarily involve unrealistically rapid changes in environmental boundary conditions, and cannot necessarily be considered as analogous to either real-world or transient future scenario conditions.

In this study we investigate the impacts of parametric uncertainty on the carbon and nitrogen responses of the CLM5. In particular, we probe a targeted parameter set comprising those parameters that determine the impacts of CO$_{2}$ and N fertilization. The full response of the model to transient climate forcing has considerably more dimensions than these two transient properties. In particular, here we have not considered responses to temperature nor to changes in the supply or demand of water. The suite of parameters that in principle control temperature responses of the model are somewhat distinct from the parameters considered here, given the large number of processes that are directly or implicitly affected by temperature (photosynthesis, respiration, soil decomposition, N mineralization, cryosphere interaction) and should be the subject of further investigations.

>>>>

\item The control state, and responses to CO$_{2}$ and nitrogen, are sensitive to parameter choice, and to baseline water and nitrogen limitation. \\

\item Parameters controlling nitrogen fixation rates dominate fertilization responses, highlighting the need for increased scrutiny of this process.  
\end{keypoints}

@@

Low values of \emph{stem\_leaf} ratio resulted in higher LAI at all sites, as well as declines in total vegetation carbon (SI Fig. 3).   The impact of increasing LAI is typically to increase light interception and hence GPP, thus these responses are typically quite large for most sites. Lower stem allocation and higher LAI also increase N demand, reducing $N_{leaf}$ and increase PNU at NWR and MET (Fig. \ref{btran state}).

Increasing $N_{costs}$ increases the respiration losses associated with nitrogen uptake.  All sites showed a moderate decline in LAI, NPP, GPP and vegetation carbon when costs were high, but lesser impacts when costs were reduced (Fig. \ref{state}). In contrast to the allocation parameters, $N_{costs}$ does not appear to be a first-order control over the pre-fertilization system state.

Increasing \emph{fracfixers} affects the maximum amount of carbon that plants can pay towards symbiotic nitrogen uptake (N fixation). For all sites and variables here, it has a relatively moderate and positive impact on the control state. As is the case for $N_{costs}$, \emph{fracfixers} does not appear to be a first-order control over the pre-fertilization system state. Interestingly, SI Fig. 1 illustrates that while \emph{fracfixers} had a large impact on the N fixation rates, this did not actually affect the overall soil Nitrogen content, implying that other factors were dominant over the equilibrium soil nutrient budget.

Changing \emph{leafcn} over one standard deviation from the mean TRY values, had substantial impacts on all the output variables at all sites. Higher \emph{leafcn} values decrease $N_{leaf}$, decreasing photosynthetic capacity and GPP.  In particular, at MET and HVF the higher \emph{leafcn} values had $N_{leaf}$ and photosynthetic capacities so low that GPP was almost zero (Fig. \ref{state}).  GPP continues to increase almost linearly with declining \emph{leafcn}  (increasing $N_{leaf}$) at all sites except HVF, indicating a lack of saturation of photosynthesis with respect to photosynthetic capacity.  This is consistent with the analysis by \cite{lawrence2018} who find that V${_c,max}$ numbers in CLM5 might be low compared to observations.  For NPP, respiration costs of modified $N_{leaf}$ and altered PNU (Figure \ref{btran state}) buffered the larger changes in GPP. At CAX, for example, the lowest \emph{leafcn} had a GPP 225 gC/m$^{2}$/y higher, while the corresponding NPP increase was only 52 gC/$m^{2}$/y.

@@

 At MET, for example, the initial GPP stimulation by CO$_{2}$ (26\%) resulted in a slow but prolonged LAI increase, (given the lower overall carbon turnover at this colder site). At 80 years, LAI had increased by 50\%, amplifying the much GPP and NPP stimulation to 53-69\% and 49-71\% respectively, (varying through the cycling climate forcing data).  For CO$_{2}$ fertilization at CAX, in contrast, the initial stimulation of GPP ($\sim$1.37) lead to an almost 50\% increase in LAI, but at this site, water limitations (lower $B_{tran,mn}$), prevented any further increase in GPP.  At HVF, the initial stimulation in the first year was lower (25\%) potentially on account of the higher carboxylation capacity at this site (HVF has a saturating response to decreasing \emph{leafcn},  Fig. \ref{state}). NWR had moderate (25\%) increases in initial GPP and LAI ($sim$30\%).  
 The stimulation of GPP via LAI here is typically greater than is typically reported in elevated CO$_{2}$ experiments, but \cite{king2005} report a large response to $sim$530 ppm CO$_{2}$ in Aspen stands, with increases in LAI and biomass of around 60\%.

 At all sites, the percentage of NPP spent on N uptake (PNU) increased, but most dramatically at MET, where initial N expenditure was lowest initially on account of low growth rates.  Further, N fixation rates increased markedly at all sites (Fig. \ref{timescales}) from their default values as shown in SI Fig 1. An increase of 2-3 was found at HVF and NWR. At MET, the very low initial rates (0.01 gNm$^{-2}$y$^{-1}$) mean that the amplification was extremely large and thus omitted from the plot.  At CAX, fixation rates initially increased and then declined through time in tandem with increases in soil Nitrogen. Soil N in general was not in equilibrium after 85 years of simulation time, but this did not appear to greatly affect the fixation rates at NWR and HVF. The soil N dynamics observed raise interesting questions about potential transient responses of CLM5 which will be covered in future analyses.

N addition most strongly affects carbon expenditure on N uptake, and thus NPP, rather than GPP directly. At NWR an initial NPP stimulation of 1.3 for NPP was magnified by subsequent increases in LAI to nearly 1.8 times the control value, causing a stimulation of 1.7 in NPP, reflected the high N limitation at this site in the pre-fertilization state. None of the other sites exhibited such a dramatic change. In particular, for CAX, which was heavily water limited in the first instance, adding nitrogen only resulting in stimulation in LAI, GPP and NPP of around 10\%. At all sites, addition of N caused a radical drop in the rates of N fixation, and gradual, non-equilibrated increases in the soil N content.

The time-dependence of the responses generates an interesting issue for interpretation. The first year responses are illustrative of the short-term immediate impacts of change, whereas the longer-term responses are heavily dominated by the feedback responses. Further, depiction of absolute vs. relative differences can serve to highlight different aspects of the model response. Here we report values from 15-20 years after fertilization began, to be consistent with the methodology used by \cite{wieder2019} in their analysis of global fertilization of the default model, and to capture the dynamics of the whole system feedbacks captured in Fig. \ref{timescales}.  Thus, figures \ref{GPP_CN} and \ref{LAI_CN} illustrate the responses to 15-20 years of CO$_{2}$ and N  fertilization simultaneously, for GPP and LAI.  Note that each figure is a different scale to best illustrate the parametric impacts.  Corresponding figures for NPP, vegetation carbon, N fixation and $N_{area}$ are in the supplemental information (SI Figures 2-5)..

@@

The impact of \emph{slatop} on fertilization was different at every site (Figs. \ref{GPP_CN} \&  \ref{LAI_CN}). For MET, \emph{slatop} was the most important control over N responses, with higher values allowing a greater (or faster) LAI/GPP feedback in response to N fertilization. At NWR, the opposite behaviour was found, where high SLA reduces the impact of N. In such cases, high pre-fertilization LAI meant that the response of GPP was more saturated, and so the feedback from increased leaf area was dampened. At HVF, higher values of \emph{slatop} slightly reduced both CO$_{2}$ and N fertilization for the same reason (although the range of variation at HVF is only 5\% between the highest and lowest values, as opposed to 40\% at NWR).

\emph{froot\_leaf} had contrasting impacts at different sites.  At MET,  low values increased the CO$_{2}$ fertilization, probably due to the increased drought stress under those conditions (Fig. \ref{btran state}). At HVF, low \emph{froot\_leaf} decreased both CO$_{2}$ and N fertilization on account of higher initial LAI, mirroring the response to \emph{slatop}.

Low \emph{stem\_leaf} ratio caused lower responses to N and CO$_{2}$ fertilization CAX and HVF for GPP and LAI at  (Figures \ref{GPP_CN} and \ref{LAI_CN}), again mirroring the NPP-LAI non-linearity issue.  At NWR,  N limitation restricted CO$_{2}$ responses, nullifying the impact of \emph{stem\_leaf}.  In contrast, MET, lower \emph{stem\_leaf} allowed greater responses to N plausibly from increasing the speed of the NPP/LAI feedback.

Higher $N_{costs}$ increase the responsiveness of the system to N at CAX, NWT and HVF.  The magnitude of the impact on N fertilization is tracks the extent of N limitation in the pre-fertilization state (Figure \ref{btran state}).   Very low values of $N_{costs}$ appeared to reduce the responsiveness of the system to CO$_{2}$ fertilization, particularly at CAX and HVF. These configurations potentially are limited in their ability to direct carbon to fixation pathways, since estimated soil N uptake costs remain low.

\emph{frac\_fixers} (the maximum fraction of NPP available for N fixation) had a particularly dominant role over the model response to step changes in N and to CO$_{2}$. For GPP and LAI, all sites displayed a 'trade-off', whereby high \emph{frac\_fixers} systems responded strongly to CO$_{2}$ but not N, and low \emph{frac\_fixers} systems responded strongly to N but not CO$_{2}$ (Fig. \ref{GPP_CN} \& \ref{LAI_CN}).   At MET,  the changing fixation capacity had no impact on the response of GPP or LAI to N fertilization.  Only \emph{slatop} and \emph{stem\_leaf} had large impacts on N fertilization for this site (which had the smallest PNU in the first instance), suggesting fundamental metabolic limitations to carbon economics and growth (e.g. a short growing season).

The impact of \emph{leafcn} on fertilization responses was not as dominant a control over fertilization responses as it was over the system state. Complex responses were illustrated, reflecting the differential costs and benefits of \emph{leafcn} on the system.  For N fertilization, at CAX and NWR, lower values of \emph{leafcn} had smaller responses to N fertilization, reflecting the higher initial LAI of these configurations and the non-linear GPP feedback discussed earlier.   At CAX, the highest \emph{leafcn} had smaller CO$_{2}$ responses. (Fig. \ref{btran state} illustrates that this configuration had less water stress, (both lower LAI and lower photosynthetic capacity should lead to smaller stomatal conductances and evapotranspiration) and thus had less water stress to be alleviated by the high CO$_{2}$.

Surprisingly, given the importance of \emph{medlyn\_slope} on stomatal responses to CO$_{2}$, we did not find a widespread impact of the Medlyn parameter on the fertilization response. Earlier studies using a less constrained range of \emph{medlyn\_slope} values indicated that it can have a potentially large impact on system state via impacts on water status, (results not shown) but that its capacity to modify the fertilization response is limited. The linear response of the stomatal conductance term to $CO_{2}$ means that the direct impact of \emph{medlyn\_slope} is likely to be restricted to changes in the pre-fertilization state.

The impact of \emph{cn\_flex\_c} was small compared to the more dominant parameters. Higher values of \emph{cn\_flex\_c} spend less carbon on N uptake when tissue N is lower than the target amount, allowing N ratios to vary more. Impacts of this parameter are manifested through their impact on $N_{leaf}$ (SI Fig. 5). For N deposition, $N_{leaf}$ always increases and for CO$_{2}$ it always declines. Smaller values of \emph{cn\_flex\_c} corresponded in all cases to smaller variations in $N_{leaf}$ under both CO$_{2}$ and N fertilization.  Lower values of \emph{cn\_flex\_c} corresponded to larger CO$_{2}$ stimulation of GPP, as these leaves maintain higher photosynthetic capacity under more N limiting conditions. The impact on NPP and LAI was less clear, given the consequence of high $N_{leaf}$ for both respiration and N uptake costs. At CAX, NWR and HVF,  the very highest values of \emph{cn\_flex\_c} \emph{also} had a higher response to CO$_{2}$ than the default. The reasons for this discontinuity are unclear, however, the impact of \emph{cn\_flex\_c} in the CLM5 code is subject to at least two 'caps', (whereby plants cannot spend less than 0.5 or more than 1.0 of the carbon that would retain the target LAI), potentially leading to complex feedbacks once these caps are reached.

\section{Discussion}
The responses of the land surface components of Earth system models to CO$_{2}$ fertilization are the subject of intensive investigations, and vast quantities of computing time will be expended on the release version of each model, under the auspices of the various CMIP6 model inter-comparison projects and associated activities (\cite{meehl2014}, \cite{eyring2016}). Throughout the CMIP activities, the carbon-cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space. Interpretation of these simulations should be informed by solid comprehension of which model properties dominate the size of the feedback strengths,  both within and between models, and ideally, a sense of how well constrained these properties are. We intend these illustrative simulations to provide initial guidance of this process. 
At the time of writing, we do not yet know the carbon-climate feedback strength of the default version of CLM5/CESM2 model. In offline simulations forced with a data atmosphere, CLM5 shows a stronger response to elevated CO$_{2}$ and a weaker response to N fertilization than previous versions of the model (Wieder et al. 2018). Further, the impact of both types of fertilization was found, in all versions of the model, to be highly variable with PFT (from 0\% to 110\% for N, -10\% to +55\% for CO$_{2}$). In this study we set out to illustrate how variations in parametric space give rise to responses to forcing. While we set out primarily to to determine responses to model parameters, findings concerning the impact of pre-fertilizaion model state, water and N limitation will have important implications for the interpretation of simulations with the default model.

@@

Aside from the straightforward direct impacts of the fixation rate on model predictions, interpretation of other configurations is less straightforward.  In some cases the sign of the impact of individual parameters changes between sites, while other parameters can demonstrate non-linear and counter-intuitive impacts as the result of numerous interacting feedbacks.  To summarize our findings, we note a variety of key outcomes of our analysis on the factors controlling responses to CO$_{2}$ and N fertilization in the CLM5, including:

@@

2. The degree of water or nutrient limitation in the initial state is also a dominant factor over the model responses to CO$_{2}$ fertilization and N deposition. Examples of this include how low \emph{froot\_leaf} increases water limitation at NWR, increasing responsiveness to CO$_{2}$ and how high \emph{leafcn} reduces water stress and thus decreases the CO$_{2}$ response at CAX.

3. Other parameters with a direct impact on fertilization responses included those related to the acquisition of nitrogen (\emph{N\_costs}) and the stoichiometric flexibility of the system (\emph{cn\_flex\_c}) illustrating the sensitivity of the model to relatively unconstrained processes within the nitrogen cycle.

4. Many of the parameters that are dominant over the initial state (\emph{slatop}, \emph{stem\_leaf} \emph{leafcn}) either have limited impacts on fertilization responses, or have impacts that are largely traceable to their impacts on the initial model state. In contrast, numerous parameters that more directly and uniformly impact fertilization responses, (\emph{frac\_fixers}, \emph{N\_costs}) are less important for the initial model state. We thus suggest that model calibration efforts focused on parameters that affect the control model state, are unlikely to constrain the major axes of responses to CO$_{2}$ and N fertilization.

5. The results can in general be interpreted as being in line with the expected dynamics of the model, given sufficient output data and consideration of the dynamics of the individual sites. Given the complexity of the model, this helps generate confidence that the model outcomes are derived from the intended process mechanisms, and are not subject to significant deviations from the intent of the model structure.

@@

\subsection{Model improvements highlighted by this analysis}
While there are many potential improvements that could be made to any land surface model, this particular assessment of the parameter space of the CLM5 in this study revealed several dynamics which highlight parts of the representation of biogeochemistry which might be improved in future versions.

Firstly, response (e.g. CAX) of NPP to increasing \emph{slatop} was often negative, despite increasing GPP. The fixed-fraction allocation scheme in the model will under some circumstances generate leaves that respire more than they assimilate under very high productivity conditions. This highlights the need for either more realistic tissue allocation, or for a scheme which optimizes leaf allocation to prevent leaf layers in negative carbon balance (\cite{fisher2010}).

@@

Thirdly, one issue requiring focus is the dominance of the 'fraction of fixers' over fertilization responses. \emph{frac\_fixers} is a proxy for the community level composition of nitrogen fixing plants. Technically, because plant density is not represented in the CLM5, this parameter represents the fraction of net assimilated carbon ($gpp$ - $r_{auto}$) that can be used for symbiotic N fixation. Ideally, \emph{frac\_fixers} would be prognostic, and the fraction of N fixing plants would respond to changes in environmental conditions and the relative competitive abilities of N fixing plants. In general, Nitrogen fixers are known to primarily exist in early-successional parts of ecosystems (\cite{vitousek1989}), and therefore representation of their ecological dynamics would likely include a representation of vegetation demographics and succession (\cite{fisher2018vegetation}, \cite{trugman2016climate}).

@@

\subsection{Applicability of results to other land surface models}
Our finding that Nitrogen fixation plays a dominant role over responses to CO$_{2}$ fertilization concurs with earlier studies. Notably, \cite{meyerholt2015} investigate six alternative models of N fixation implemented within the O-CN model (\cite{zaehle2010}). They conclude that their optimizing algorithm, which, as in CLM5, switches N sources depending on their relative costs, represents a promising means by which N fixation might be simulated. Interestingly, this model also includes a maximum rate of fixation per unit (root) biomass parameter, but analysis to it's sensitivity is not shown. Further, given the difficulty of matching observations of complex N cycle processes with more abstract 'big leaf' models such as CLM5, and the ecological complexities governing the abundance their abundance, \cite{meyerholt2015} also advocate for the dynamic simulation of the distribution of N fixing plants within ESMs, as opposed to the assumption that maximum N fixing capacity if static in space and time.

Given that the response of GPP is by definition non-linear with respect to LAI, (given the saturating properties of light absorbtion in increasingly dense canopies) we anticipate that the impacts of initial state will be similar in most land surface models. While it is perhaps not a novel point, water limitation should act to modulate CO$_{2}$ fertilization, while N fertilization will of course be stronger in areas that are more N limited in the first instance.

@@

\subsection{Limitations and future directions}
In this study we investigate the impacts of parametric uncertainty on the carbon and nitrogen responses of the CLM5. In particular, we probe a targeted parameter set comprising those parameters that determine the impacts of CO$_{2}$ and N fertilization. The full response of the model to transient climate forcing has considerably more dimensions than these two transient properties. In particular, here we have not considered responses to temperature nor to changes in the supply or demand of water. The suite of parameters that in principle control temperature responses of the model are somewhat distinct from the parameters considered here, given the large number of processes that are directly or implicitly affected by temperature (photosynthesis, respiration, soil decomposition, N mineralization, cryosphere interaction) and should be the subject of further investigations.

Many kinds of further analyses are of course possible, including ramping transient simulations,  global simulations with alternative parameterizations, calibration of critical parameters against manipulation experiments, wider assessment against greater ranges of flux tower data, and tighter constraints of the parameter ranges used. We hope that this initial assessment of the CLM5 parameter space will provide a starting point for such activities.
